---
title: "ddos logs arff format test"
output: html_notebook
---

Data from: https://www.kaggle.com/jacobvs/ddos-attack-network-logs

> The data was collected from a larger-scale network and thus depending on your usage you can ignore a lot of the first attributes of the dataset as they're more network specific.

##Context
The dataset contains around 2,100,000 labelled network logs from various types of network attacks.
The types of network attacks logged are: UDP-Flood, Smurf, SIDDOS, HTTP-FLOOD, & Normal traffic

##Content
Each line of data contains these attributes:

* @attribute SRC_ADD numeric
* @attribute DES_ADD numeric
* @attribute PKT_ID numeric
* @attribute FROM_NODE numeric
* @attribute TO_NODE numeric
* @attribute PKT_TYPE {tcp,ack,cbr,ping}
* @attribute PKT_SIZE numeric
* @attribute FLAGS {-------,---A---}
* @attribute FID numeric
* @attribute SEQ_NUMBER numeric
* @attribute NUMBEROFPKT numeric
* @attribute NUMBEROFBYTE numeric
* @attribute NODENAMEFROM {Switch1,Router,server1,router,clien-4,client-2,Switch2,client-5,clien-9,clien-2,clien-1,clien-14,clien-5,clien-11,clien-13,clien-0,switch1,client-4,clienthttp,clien-7,clien-19,client-14,clien-12,clien-8,clien-15,webserverlistin,client-18,client-1,switch2,clien-6,client-10,client-7,webcache,clien-10,client-15,clien-3,client-17,client-16,clien-17,clien-18,client-12,client-8,client-0,clien-16,client-13,client-11,client-6,client-3,client-9,client-19,http_client}
* @attribute NODENAMETO {Router,server1,Switch2,Switch1,clien-1,clien-5,clien-7,switch1,clien-11,clien-15,clien-13,clien-3,clien-9,clien-6,router,clien-4,clien-14,switch2,clien-8,clienthttp,webcache,clien-10,clien-12,webserverlistin,clien-0,clien-2,http_client,client-13,client-9,client-1,client-19,client-4,client-17,client-7,client-3,client-12,client-2,clien-18,client-16,clien-17,client-0,clien-16,client-18,client-5,client-11,client-14,client-8,client-6,client-10,clien-19,client-15}
* @attribute PKT_IN numeric
* @attribute PKT_OUT numeric
* @attribute PKT_R numeric
* @attribute PKTDELAYNODE numeric
* @attribute PKT_RATE numeric
* @attribute BYTE_RATE numeric
* @attribute PKTAVGSIZE numeric
* @attribute UTILIZATION numeric
* @attribute PKT_DELAY numeric
* @attribute PKTSENDTIME numeric
* @attribute PKTRESEVEDTIME numeric
* @attribute FIRSTPKTSENT numeric
* @attribute LASTPKTRESEVED numeric
* @attribute PKT_CLASS {Normal,UDP-Flood,Smurf,SIDDOS,HTTP-FLOOD}

```{r}
require("farff")
require("tidyverse")
require("caret")
require("nnet")


d = readARFF("final-dataset.arff") #this takes a few seconds to load

#d = readARFF("final-dataset-short.arff") #this is the short dataset that I am using only to do some quick initial analysis

summary(d)
typeof(d)
dim(d)
class(d)
```
No missing values!
There are some categorical variables, including the target variable: PKT_TYPE, FLAGS, NODE_NAME_FROM and NODE_NAME_TO, 

# Target variable
PKT_CLASS is the target variable, but it's a discret categorical variable with 5 possible values: Normal,UDP-Flood,Smurf,SIDDOS,HTTP-FLOOD. I'll create 4 dummy variables with binary values (1 or 0) so I can use regression to predict each one of them:
* udp 
* smurf
* siddos
* http

the variable "left out" (reference group) is "Normal".

```{r}
contrasts(d$PKT_CLASS)
res <- model.matrix(~ PKT_CLASS, data = d)
head(res[, -1])

pkt_class_dummies <- data.frame(res[, -1])
names(pkt_class_dummies)
dim(pkt_class_dummies)
colnames(pkt_class_dummies) = c("udp_flood", "smurf", "siddos", "http_flood")
```

....should i create just one dummy var for normal or not (ddos attack)?

```{r}
d_augmt <- d
d_augmt$PKT_CLASS_UDP_FLOOD <- pkt_class_dummies$udp_flood
d_augmt$PKT_CLASS_Smurf<- pkt_class_dummies$smurf
d_augmt$PKT_CLASS_SIDDOS<- pkt_class_dummies$siddos
d_augmt$PKT_CLASS_HTTP_FLOOD<- pkt_class_dummies$http_flood

summary(d_augmt)
```

## Splitting the data into 3 datasets
Training,
Vaidation
Test (reserved for testing final models)

```{r}
set.seed(123) #for reproducibility
ss <- sample(1:3,size=nrow(d_augmt),replace=TRUE,prob=c(0.6,0.2,0.2))
train_data <- d_augmt[ss==1,]
val_data <- d_augmt[ss==2,]
holdout_data <- d_augmt[ss==3,]
```
## correlations

```{r}
corr_matrix <- cor(d[, sapply(d, is.numeric)]) #removing non numeric indexes
library("corrplot")
corrplot(corr_matrix)
#or heatmap version of correlogram:
palette = colorRampPalette(c("green", "white", "red")) (20)
heatmap(x = corr_matrix, col = palette, symm = TRUE)
```
## fitting a quick linear model
```{r}
model <- lm(PKT_CLASS_UDP_FLOOD ~ PKT_AVG_SIZE, data = train_data)
anova(model)
plot(model$fit)
```

##Data normalization

##Feature selection - load and transform bigger dataset first (training)
###Correlations